---
title: "Lab 8 - Vector-borne disease models"
author: "sbsambado"
date: "4/4/2020"
output: html_document
---

Continuous-time SI models that incorporate disease vectors (2 different species)

Issues that need to be considered with modeling
1. Natural history
-biology of host, vector, and their interactions
2. Time period of the model
-births and deaths of vectors often icorporated beacuse they have short generation times cause vs. host
3. Research question

Basic vector SIS model
4 state variables
1. susceptible host (S)
2. susceptible vector (U)
3. infected host (I)
4. infected vector (V)

5 parameters
1. inoculation rate (beta)
2. acquisition rate (alpha)
3. host turnover rate (gamma)
4. vector birth rate (b_y)
5. vector death rate (mu_v)

Numerical simulations of basic SIS model
```{r}
#library(deSolve)

SISvector <- function(dt, inits, parms) {
  with(as.list(c(parms, inits)), {
    dS = - beta*S*V/ (S + I) + gamma*I
    dI =  beta*S*V/ (S + I) - gamma*I
    dU = -alpha*U*I/ (S + I) + b*V - mu*U
    dV = alpha*U*I/ (S + I) + b*V - mu*V
    return(list(c(dS, dI, dU, dV)))
  })
}

# set time
dt <- seq(0, 150, by = 1)

# set initial values
S0 <- 99
I0 <- 1
U0 <- 200
V0 <- 10
inits <- c(S = S0, I = I0, U = U0, V = V0)

alpha = 0.5
beta = 0.6
gamma = 0.5
mu = 0.15
b = 0.1

parms <- c(alpha = alpha, beta = beta, gamma = gamma, mu = mu, b = b)

# numerical situation
model.out <- as.data.frame(ode(y = inits, times = dt, func = SISvector, parms = parms))
tail(model.out)

matplot(model.out[,-1],
        type = 'l', xlab = 'time', ylab = 'population',
        col = c('green', 'green', 'blue', 'blue'))
```

Investigate the effect of vector population growth

What are the effects of vector population growth on disease prevalence?
- calculate infected host density (I) over range of vector birth rates over a long time period
```{r}
# make a vector of values for vector birth rate parameter 'b' and extend time

bVec <- seq(0, 0.2, by = 0.01)
dt <- seq(0, 500, by = 1)

SISVectorFunc <- function(x) { # run varibale x to the params vector to represent birth rate values
  parms <- c(alpha = alpha, beta = beta, gamma = gamma, mu = mu, b = b)
  model.out2 <- as.data.frame(ode(y = inits, times = dt, func = SISvector,  parms = parms))
  infectFinal <- (model.out2$I[501]/ (S0 + I0))* 100 # % of host infected
  return(c(x, infectFinal))
}

bResults <- as.data.frame(t(sapply(bVec, SISVectorFunc, simplify = TRUE)))
names(bResults) <- c('b', 'I')

plot(I ~ b, data = bResults, type = 'b', lwd = 2, col = 'green',
     ylab = '% of Infected hosts', xlab = 'vector birth rate')
```

