---
title: "Lab 5 & 6 - Loops, functions, chain binomial"
author: "sbsambado"
date: "3/27/2020"
output: html_document
---

BIOL 490 is a class taught by Dr. Andrea Swei, that I was a teaching assistant for

Two types of loops:
1. for-loop: runs a process several times for X number of times, when it's done it exits the loop and continues on to the rest of the program

2. while-loops: repeats some function (loop) until a defined criteria or condition is met. A while loop specify that as long as a parameter x is less than some value, the function will continue to loop.

for-loop examples
```{r}
Popn <- 1 # create initial value
Gen <- 10 # number of generations

for( n in 2:Gen) {
  Popn[n] <- Popn[n-1] * 2 # create vector to put values from 2 to Gen
}
Popn

plot(1:Gen, Popn, type = 'b')

# Example 2

# Calculate the circumference of a circle using value of previous time point

circ <- 1 

for(n in 2:Gen) {
  circ[n] <- 3.14*circ[n-1]*2
}
circ

plot(1:Gen, circ, type = 'b')

# Exercise 3

# Calculate the square of the Popn vector
Popn.squared = NULL

for(n in 1:Gen) {
  Popn.squared[n] <- Popn[n]^2
}
Popn.squared

plot(1:Gen, Popn.squared, type = 'b')

# Exercise 4

# pop starts at 1 and grows at a rate of 40% per year for 20 generations
pop <- 1
gen <- 20
for(n in 2:gen) {
  pop[n] <- pop[n-1]*1.4
}
pop

plot(1:gen, pop, type = 'b')
```


While-loop examples
```{r}
# Example 1

pop.vec = pop.now <- 1
count <- 1 # iterations

while(pop.now <=25000) {
  pop.now <- pop.now*2 # growth rate for each generation
  pop.vec <- c(pop.vec, pop.now) # keeps adding on the most recent pop.now
  count <- count + 1 # keeps track of number of gererations to exceed 2500
}
pop.vec

# modify code to see how many cycles to exceed 1 million
pop.vec = pop.now <- 1
count <- 1 # iterations

while(pop.now <=1000000) {
  pop.now <- pop.now*2 # growth rate for each generation
  pop.vec <- c(pop.vec, pop.now) # keeps adding on the most recent pop.now
  count <- count + 1 # keeps track of number of gererations to exceed 2500
}
pop.vec
count # how many cycles to reach 1 million

```

Creating functions
```{r}
# myfx <- function(x,y,z) { # arguments
#   expression_1 # what function will do with arguments
#   expression_2
#   expression_3
#   Output # if you want to output automatically
# }

# function to calculate mean
ave <- function(x) {
  mu <- sum(x)/length(x)
  mu
}

# now create a vector to run function with
vec <- c(2,8,10,3,7,8,4,23)

ave(vec)
mean(vec) # should match ave(vec)
```

Write a function to produce 95% Credible intervals assuming a normal distribution

CI at 95% level assume that 2.5% of curve is on either side of the distribution of the mean and is calculated as mean +/- 1.96x standard error (se) of the effect size
```{r}
CI95 <- function(x) {
  mean <- mean(x)
  sd <- sd(x)
  n <- length(x)
  se <- sd/sqrt(n)
  
  CI.vec <- c((mean-(1.96*se)), mean, (mean + (1.96*se)))
  names(CI.vec) <- c('2.5%', 'Mean', '97.5%')
  CI.vec
}

fake_data <- c(10,3,6,8,9,8,54,5,2,3,43,25,2,5,2,45,66,78,9,80,345,56,7,8,9)

CI95(fake_data)
```

# The Chain Binomial Model
Chain Binomial model is based on the assumption of discrete generations and is basically a difference equation model
- population grows in steps and not continuously
- not very applicable for many things, but maybe insects that molt or species with 1 reproductive cycle per year
- but, a model should only be as complicated as it needs to be

chain binomial model (cbm) is a simple stochastic model with discrete generations
- each generation, new infections are distributed accroding to thee binomial distribution 
- binomial distribution describes the probaility distribution function on an event that has only  1 of 2 outcomes, thus binomial

The number of new infections, I, is binomially distributed
The number of trials is equal to number of susceptibles, S_t, and the probability of infectio, p = 1 - exp(-beta*I_t)
```{r}
# create a chain binomial function

# one step model
chain.binomial.onestep <- function(x, params) {
  S <- x[1]
  I <- x[2]
  beta <- params['beta']
  new.I <- rbinom(n =1, size = S, prob=1-exp(-beta*I))
  new.S <- S - new.I
  c(S = new.S, I = new.I)
  
}


# one step not practical, now make a multi-step model

chain.binomial.model <- function(x, params, nstep) {  
  X <- array(dim = c(nstep+1, 3))
  colnames(X) <- c('time', 'S', 'I')
  X[1,1] <- 0
  X[1, -1] <- x
  
  for(k in 1:nstep) {
    X[k+1, 1] <- k
    X[k+1, -1] <- x <- chain.binomial.onestep(x, params)
  }
      X
}

# initial values
set.seed(38499583)
nsims <- 10
nstep <- 20
xstart <- c(S = 200, I = 5)
params <- c(beta = 0.1)
x <- vector(mode = 'list', length = nsims)
for (k in 1:nsims) {
  x[[k]] <- as.data.frame(chain.binomial.model(xstart, params, nstep))
}
plot(c(0,6), c(0,200), type = 'n', xlab = 'generation', ylab = 'incidence')
for(k in 1:nsims) {
  lines( I ~ time, data = x[[k]], col = k, type = 'o')
}


```

