---
title: "Lab 12 - Macroparasite aggregation"
author: "sbsambado"
date: "4/4/2020"
output: html_document
---

Why are macroparasites aggregated among hosts?

This question comes down to differences in the transmission rate of the pathogen. In a regular SIR model, the rate that a given susceptible individual becomes infected is beta * I

If every host has the same probability of becoming infected, then beta is the same for all hosts. For instance, everyone could have a 10% chance of becoming infected at any particular moment.This would follow a poisson distribution.

But, beta, is actually a product of the average per capita contact rate between individuals (c) and the average probability of successful transmission given contact (v)
beta = c * v

Generate a poisson distribution based on parameter values (c and v). In order to generate a poisson distribution, you need to know the mean in a poisson distribution (aka lambda) is determined by a rate parameter, r, multiplied by the length of time that the process has been occuring, t.
- In this situation the rate in question is the transmission rate, beta, which we have defined as beta = c * v
- Therefore, lambda is defined by beta *t (rate = beta = c*v)
```{r}
# assume contact rate is 50%
c = 0.5

# probability of transmission if there is contact is 20%
v = 0.2

# lambda
t = 20
lambda = (c*v)*t

help(rpois)

# create a vector of numbers drawn from a poisson distribution that represents the number of parasites in each individual host for 1000 hosts
parasites <- rpois(n = 1000, lambda = lambda)

# quantify whether the occurrences are clustered together or dispersed (is there high variance in how many parasites a host has, or do most hosts have ~same number of parasites?)
var(parasites)/mean(parasites)

hist(parasites, right = FALSE, ylab = 'number of hosts with X parasites',
     xlab = 'number of parasites per host')
```

Above chunk assumed that the chance of becoming infected was exactly the same between everybody. But what if individuals had different probabilities of getting infected? Beta varies among hosts?
- Poisson distribution is not right distribution then. 
- Compound distribution where the infection is still a random process but now the rate parameter (B*t = lambda) is no longer constant. Instead, the rate parameter must also vary in a random way
- now rate parameter is a drawn from gamma distribution
```{r}
rate <- rgamma(10000, shape = 1, scale = 10)
parasitesA <- rpois(n = 10000, lambda = rate)

hist(parasitesA, right = FALSE, ylab = 'number of hosts with X parasites')

# variance to mean ratio
var(parasitesA)/mean(parasitesA)
```

Does the 80/20 rule hold up? Sort parasites from highest to lowest number
```{r}
parasitesB <- sort(parasitesA, decreasing = T)

# 20% of individuals out of 10000 is 2000, calculate sum of first 2000
aggregates <- sum(parasitesB[1:2000])

heavily_tot <-aggregates/10000

```

